---
title: "HW6_Mkd"
author: "Chad Melton and Rebecca Brink"
date: "11/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Agresti Chapter 6: 6.1, 6.6, 6.8, 6.12, 6.16, 6.17, 6.22
```{r}



library(MASS)
library(rpart)
library(epiR)
library(ggplot2)
library(vcdExtra)
library(vcd)
library(lmtest)
library(caret)
library(ROCR)
library(pROC)
library(VGAM)
```
# 6.1
' A model fit predicting preference for President (Democrat, Republican, Independent)
using x = annual income (in $10,000 dollars) is log(PiD/PiI ) = 3.3 Pi? 0.2x and log(PiR/PiI ) = 1.0 + 0.3x.'
\
\
a) State the prediction equation for log(PiR/PiD). Interpret its slope.

log(PiD/PiI ) = 3.3 Pi? 0.2x and log(PiR/PiI ) = 1.0 + 0.3x.

THe prediction equation is log(PiR/PiI )-log(PiD/PiI)
=log(PiR)-log(PiD)=-2.3 +5x
Now if x = 10000, the slope suggests that republicans increase by 0.5 per $10k.
\
\
b) Find the range of x for which PiR > PiD.
So for PiR > PiD, after a few manipulations we get to exp(1.0 + 0.3x) >exp(3.3-0.2x)->
0.5x > 2.3-> x > 4.6. Therefore, for x = $10000, the range is greater than $46000 for 
republicans.'

c) State the prediction equation for PiI.
 PiI= 1/(exp(1+0.3x) + exp(3.3-0.2x))'

# 6.6 
Does marital happiness depend on family income? For the 2002 General Social
Survey, counts in the happiness categories (not, pretty, very) were (6, 43, 75)
for below average income, (6, 113, 178) for average income, and (6, 57, 117)
for above average income. Table 6.15 shows output for a baseline-category
logit model with very happy as the baseline category and scores {1, 2, 3} for
the income categories.

a) Report the prediction equations from this table.

log(Pi1/Pi3) = -2.5551 - 0.2275x
log(Pi2/Pi4) = -0.3513 - 0.0962x

b) Interpret the income effect in the first equation.
Here, the income and happiness both increase so it appears that there could
be some relationship here. However, it should be mentioned that the P-Values are all >> than
0.05 except for the intercept.

c) Report theWald test statistic and P-value for testing that marital happiness
is independent of family income. Interpret.
The given Wald Test statistic is reported to be 0.943 and the P-Value is -0.6240.
These results suggest that you dont need to be rich to have a happy marriage (although
it would be nice).


d) Does the model fit adequately? Justify your answer.The ratio of the deviance/df suggests
that the model is not a great fit. The provided pvalue suggests the same.


e) Estimate the probability that a person with average family income reports
a very happy marriage.

Pi3 = 1/(1+exp(-2.5551-2 x 0.2275) + exp(-0.3513-2 x 0.0962)))
   = 1/1.629
  =0.61
   '
   
# 6.8
Table 6.17 results from a clinical trial for the treatment of small-cell lung cancer.
Patients were randomly assigned to two treatment groups. The sequential therapy
administered the same combination of chemotherapeutic agents in each
treatment cycle. The alternating therapy used three different combinations,
alternating from cycle to cycle.


a. Fit a cumulative logit model with main effects for treatment and gender.
Interpret the estimated treatment effect.'

```{r}


small.cell = read.csv("SmallCell.csv", header=TRUE)
small.cell$Response = ordered(small.cell$Response,c("Pro","NoCh", "ParRem", "CompRem"))

(fit.small.cell <- vglm(Response~small.cell$Gender+small.cell$Therapy, cumulative(parallel= TRUE, reverse = FALSE), weights = Count, data=small.cell))
summary(fit.small.cell)
exp(-0.5807)
exp(-0.5414)
```
For treatment the coefficient is -0.5807 and calculates to odds of 0.5595 for sequential therapy vs alternating. 
For gender, the coefficient is -0.5414 to which calculates to odds if 0.581 for males response vs female. 

b. Fit the model that also contains an interaction term between treatment
and gender. Interpret the interaction term by showing how the estimated
treatment effect varies by gender.
```{r}





(fit.small.cell2 <- vglm(Response~small.cell$Gender+small.cell$Therapy +(small.cell$Therapy*small.cell$Gender), cumulative(parallel= TRUE, reverse = FALSE), weights = Count, data=small.cell))
summary(fit.small.cell2)


exp(0.5904)
```

The interaction term between gender and therapy is 0.5904 with odds of ~1.8.'

c. Does the interaction model give a significantly better fit?
By comparison of the resisidual deviance to DF ratio of both models. We see that the deviance is 788 for the first model and 789 for the interaction model. These values suggest that there is not really a significant difference in the two models.


# 6.12
Table 6.18 shows results from the 2000 General Social Survey relating happiness
and religious attendance (1 = at most several times a year, 2 = once a
month to several times a year, 3 = every week to several times a week).

```{r}
happy = read.csv("Happy612.csv")
happy$Happiness = ordered(happy$Happiness, c("NTH", "PH", "VH"))
fit.happy = vglm(Happiness~happy$Religion, cumulative(parallel = TRUE, reverse = FALSE),
                 weights = Count, data = happy )
summary(fit.happy)

```
a. Fit a multinomial model. Conduct descriptive and inferential analyses about
the association.

```{r}

happy2 = reshape(happy[,c("Religion", "Happiness", "Count")], timevar = c("Happiness"),idvar = c("Religion"), direction = "wide" )
fit.happy2 = vglm(cbind(Count.NTH, Count.PH, Count.VH)~Religion, cumulative(parallel = TRUE, reverse = FALSE), data = happy2)
summary(fit.happy2)

```
In this situation, because the religious coefficient is negative, as x increase, the probability 
decreases. The summary also provides a very small p-value indicating that we shall reject the null hypothesis. '

b. Analyze the model goodness of fit.
The residual deviance to DF ratio suggests that this model fits our data well.








